---
title: "Causal effect of having more bronze or silver medals on placement"
format: html
author: "Hunter Stephens"
execute: 
  echo: False
---

```{r}
#| label: setup
#| message: False
#| warning: False
library(tidyverse)
library(tidybayes)
library(brms)
library(gtsummary)
```

```{r}
#| label: cleaning
#| message: False
summer_med <- read_csv("Summer_olympic_Medals.csv")
sorted_summer_med <- summer_med |>
  arrange(Year, 
          desc(Gold), 
          desc(Silver), 
          desc(Bronze)) |>
  group_by(Year) |>
  mutate(placement = row_number()) |> 
  mutate(medal_type = case_when(
    Silver > Bronze ~ "More Silver",
    Bronze > Silver ~ "More Bronze",
    TRUE ~ "Equal"
  )) |>
  ungroup() |> 
  drop_na()

```

```{r}
#| label: model
#| cache: True
fit_summer <- sorted_summer_med |> 
    brm(formula = placement ~ Silver + Bronze, family = poisson(), silent = 2, refresh = 0, seed = 9)
ndata <- expand.grid(
  Silver = 0:50, 
  Bronze = 0:50
)

# Ensure the data is expanded correctly
epred_draws <- fit_summer |> 
  add_epred_draws(newdata = ndata)

graph <- epred_draws |> 
  mutate(medal_type = case_when(
    Silver > Bronze ~ "More Silver",
    Bronze > Silver ~ "More Bronze",
    TRUE ~ "Equal"
  )) |>
  sample_frac(.1) 
```

The number of gold medals directly effects olympic placings however, how do bronze and silver medals effect it? We gathered a dataset off of Kaggle ([Data](https://www.kaggle.com/datasets/ramontanoeiro/summer-olympic-medals-1986-2020) containing summer olympic placings between the years 1896 - 2020. using this we seek to find the effect of having more silver or bronze medals on scoring. We modeled the amount bronze and silver metals as covariates to placement, our outcome, then mutated the data to find which there were more of. We then saved this variable as the treatment. We found that Silver medals have a much stronger effect on placing than bronze medals

```{r}
ggplot(graph, aes(x = ifelse(medal_type == "More Silver", Silver, Bronze), y = .epred, color = medal_type)) +
  geom_jitter(alpha = 0.01) +
  labs(
    title = "Posterior Predictive Distribution of Placements by Medal Type",
    subtitle = "Silver has a much stronger correlation to placement than bronze",
    x = "Number of Medals",
    y = "Predicted Placements",
    color = "Medal Type"
  ) +
  scale_color_manual(values = c("More Silver" = "grey", "More Bronze" = "tan"), guide = guide_legend(override.aes = list(alpha = 1))) +
  scale_y_reverse() +
  theme_minimal()
```

 
